---
title: 'Hepatoprotective effects of systemic ER activation'
subtitle: 'ChIPseq/Epigenome genome - Quantification of reads in diffbound promoters and enhancers'
author: 'Christian Sommerauer & Carlos Gallardo'
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    highlight: tango
---

```{r setup, include=FALSE}
# set path
path <- strsplit(dirname(rstudioapi::getActiveDocumentContext()$path), '/')[[1]]
path <- path[-c((length(path)-1):length(path))]
path <- paste(path, collapse = '/')
knitr::opts_knit$set(root.dir = path)
knitr::opts_chunk$set(echo = T, message = F, warning = F, dev = 'png')
```

# This script has two parts: 

### in Part 1, we use thr generated bed file with 182 diffbound promoters (one TSS per gene; the file after diffbind analysis still has multiple TSS per gene, 184 total). This bedfile is then converted into a SAF file using awk in command line. This SAF file, together with BAM files is used for featurecounts to generate a count matrix with all reads in these peak regions (this is analyzed on UPPMAX). The same is done for the 1,816 enhancer peaks.

### in Part 2, we import this count matrix into R again, normalize it, calculate significances and generate the plots.


# Part 1
```{bash, eval = FALSE}
#The enhancer and promoter (with unique TSS, generated in the chunk above) bed files were used to create a saf file using the following command:

awk 'OFS="\t" {print $1"."$2"."$3, $1, $2, $3, "."}' in.bed > out.saf

#The saf file was then used to annotate the reads, all bam files (Remove duplicates, blocklist, sorted,..) were used.
#subread version 2.0.0 was used. 

Following featureCounts chunk was used:
featureCounts \
        -g gene_id \
        -s 2 \
        -T 2 \
        -M \
        -F SAF \
        -O \
        -C \
        -a ${SAF_PATH}/example.saf  \
        -o ${OUTPUT_PATH}/example.readCount \
        ${BAM_PATH}/*H3K27ac*MkDup.bam \
        &> ${OUTPUT_PATH}/example.readCount.log

the code was then uploaded to uppmax, features counted, then downloaded, finally the header cleaned up (using the bash clean header script).
```


# Part 2

### Import the reads in peaks which was created using feature counts. Then, normalize the reads to do the analysis. Perform an Anova to compare significances between the conditions. 

#### First, for promoters.
```{r}
library(tidyverse)
counts_prom <- read.delim("results/Epigenome_analysis/DAc_promoters_142_H3K27ac.clean.readCount", header=T) %>% tibble::column_to_rownames("Geneid") %>%
  select("CK0744_H3K27ac_CD2", "CD6_H3K27ac_S2_R1_001", "CK0745_H3K27ac_CD9",
         "CK0746_H3K27ac_HFD3", "CK0747_H3K27ac_HFD4", "HFD6_H3K27ac_S8_R1_001", 
         "CK0748_H3K27ac_DPN2", "CK0749_H3K27ac_DPN3", "DPN6_H3K27ac_S15_R1_001",
         "DIP3_H3K27ac_S9_R1_001",  "DIP6_H3K27ac_S10_R1_001", "DIP10_H3K27ac_S11_R1_001",
         "E2_2_H3K27ac_S16_R1_001", "CK0750_H3K27ac_E2_8", "CK0751_H3K27ac_E2_9",
         "PPT1_H3K27ac_S12_R1_001",  "PPT2_H3K27ac_S13_R1_001", "PPT3_H3K27ac_S14_R1_001")

names(counts_prom) <- c("CDm2","CDm6", "CDm9","HFDm3","HFDm4","HFDm6", "DPN2","DPN3","DPN6","DIP3","DIP6","DIP10","E2_2", "E2_8","E2_9", "PPT1","PPT2","PPT3")
colsums_prom <- colSums(counts_prom[,])

#normalise per depth
counts_prom_norm <- sweep(counts_prom, 2, colsums_prom, FUN = "/") 
counts_prom_norm2 <- counts_prom_norm *10^6
colSums(counts_prom_norm2[,])

#Combine the replicates
counts_prom_norm3 <- counts_prom_norm2 %>% 
  mutate(avg_CD = rowMeans(counts_prom_norm2[,1:3])) %>%
  mutate(avg_HFD = rowMeans(counts_prom_norm2[,4:6])) %>%
  mutate(avg_DPN = rowMeans(counts_prom_norm2[,7:9])) %>%
  mutate(avg_DIP = rowMeans(counts_prom_norm2[,10:12])) %>%
  mutate(avg_E2 = rowMeans(counts_prom_norm2[,13:15])) %>%
  mutate(avg_PPT = rowMeans(counts_prom_norm2[,16:18])) %>%
  tibble::rownames_to_column("loc_ID")

prom_enh_diffbound <- readRDS("results/Epigenome_analysis/annotated_diffbind_and_genomewide_promoters_enhancers.rds")

Prom_Diffbind_HFDup <- prom_enh_diffbound$promoters_HFDup %>% mutate("loc_ID" = paste0(seqnames, ".",start,".",end), .before = seqnames)
Prom_Diffbind_HFDdown <- prom_enh_diffbound$promoters_HFDdown %>% mutate("loc_ID" = paste0(seqnames, ".",start,".",end), .before = seqnames)

counts_prom_HFDup <- counts_prom_norm3 %>% filter(loc_ID%in%Prom_Diffbind_HFDup$loc_ID) %>% dplyr::select(1,20:25) 


counts_prom_HFDup2 <- counts_prom_HFDup %>%
  pivot_longer(cols=2:7) %>% mutate("updown"= "HFDup") %>% group_by(name)

# STATISTICS - One-sided anova.
counts_prom_HFDup.stat <- aov(value ~ name, data = counts_prom_HFDup2)
summary(counts_prom_HFDup.stat)
counts_prom_HFDup.stat1 <- TukeyHSD(counts_prom_HFDup.stat)
counts_prom_HFDup.stat2 <- counts_prom_HFDup.stat1$name

counts_prom_HFDdown <- counts_prom_norm3 %>% filter(loc_ID%in%Prom_Diffbind_HFDdown$loc_ID) %>% dplyr::select(1,20:25)
counts_prom_HFDdown2 <- counts_prom_HFDdown %>%
  pivot_longer(cols=2:7) %>% mutate("updown"="HFDdown") %>% group_by(name)

# STATISTICS - One-sided anova.
counts_prom_HFDdown.stat <- aov(value ~ name, data = counts_prom_HFDdown2)
summary(counts_prom_HFDdown.stat)
counts_prom_HFDdown.stat1 <- TukeyHSD(counts_prom_HFDdown.stat)
counts_prom_HFDdown.stat2 <- counts_prom_HFDdown.stat1$name

counts_prom_HFD_plot <- rbind(counts_prom_HFDup2, counts_prom_HFDdown2)
counts_prom_HFD_plot$value <- counts_prom_HFD_plot$value+1
```

#### Plot for promoters
```{r}
order <- c("avg_CD", "avg_HFD", "avg_DPN","avg_DIP", "avg_E2", "avg_PPT")
ggplot(counts_prom_HFD_plot, aes(x=factor(name, levels=order), y=log2(value), fill=updown)) +
  geom_boxplot(alpha=0.8) +
  theme_classic() +
  facet_wrap(vars(updown)) + 
  xlab("") +
  ggtitle("Promoters") +
  ylab("H3K27ac signal\n log2(normalized read count in \nDAc promoter peaks in CDvsHFD)") +
  scale_fill_manual(values = c("#88CCEE",  "#332288", "#DDCC77", "#CC6677", "#AA4499",  "#882255"))
```

#### Then, for enhancers
```{r}
library(dplyr)
library(tidyr)
counts_enha <- read.delim("results/Epigenome_analysis/DAc_enhancers_2181_H3K27ac.clean.readCount", header=T) %>% tibble::column_to_rownames("Geneid") %>%
  select("CK0744_H3K27ac_CD2", "CD6_H3K27ac_S2_R1_001", "CK0745_H3K27ac_CD9",
         "CK0746_H3K27ac_HFD3", "CK0747_H3K27ac_HFD4", "HFD6_H3K27ac_S8_R1_001", 
         "CK0748_H3K27ac_DPN2", "CK0749_H3K27ac_DPN3", "DPN6_H3K27ac_S15_R1_001",
         "DIP3_H3K27ac_S9_R1_001",  "DIP6_H3K27ac_S10_R1_001", "DIP10_H3K27ac_S11_R1_001",
         "E2_2_H3K27ac_S16_R1_001", "CK0750_H3K27ac_E2_8", "CK0751_H3K27ac_E2_9",
         "PPT1_H3K27ac_S12_R1_001",  "PPT2_H3K27ac_S13_R1_001", "PPT3_H3K27ac_S14_R1_001")

names(counts_enha) <-  c("CDm2","CDm6", "CDm9","HFDm3","HFDm4","HFDm6", "DPN2","DPN3","DPN6","DIP3","DIP6","DIP10","E2_2", "E2_8","E2_9", "PPT1","PPT2","PPT3")
colsums_enha <- colSums(counts_enha[,])

counts_enha_norm <- sweep(counts_enha, 2, colsums_enha, FUN = "/") 
counts_enha_norm2 <- counts_enha_norm *10^6
colSums(counts_enha_norm2[,])

#Combine the replicates
counts_enha_norm3 <- counts_enha_norm2 %>% 
  mutate(avg_CD = rowMeans(counts_enha_norm2[,1:3])) %>%
  mutate(avg_HFD = rowMeans(counts_enha_norm2[,4:6])) %>%
  mutate(avg_DPN = rowMeans(counts_enha_norm2[,7:9])) %>%
  mutate(avg_DIP = rowMeans(counts_enha_norm2[,10:12])) %>%
  mutate(avg_E2 = rowMeans(counts_enha_norm2[,13:15])) %>%
  mutate(avg_PPT = rowMeans(counts_enha_norm2[,15:18])) %>%
  tibble::rownames_to_column("loc_ID")

Enha_Diffbind_HFDup <- prom_enh_diffbound$enhancers_HFDup %>% mutate("loc_ID" = paste0(seqnames, ".",start,".",end), .before = seqnames)
Enha_Diffbind_HFDdown <- prom_enh_diffbound$enhancers_HFDdown %>% mutate("loc_ID" = paste0(seqnames, ".",start,".",end), .before = seqnames)

counts_enha_HFDup <- counts_enha_norm3 %>% filter(loc_ID%in%Enha_Diffbind_HFDup$loc_ID) %>% dplyr::select(1,20:25) 
counts_enha_HFDup2 <- counts_enha_HFDup %>% 
  pivot_longer(cols=2:7) %>% mutate("updown"= "HFDup")

# STATISTICS - One-sided anova.
res.aov.enha.HFDup.stat <- aov(value ~ name, data = counts_enha_HFDup2)
summary(res.aov.enha.HFDup.stat)
res.aov.enha.HFDup.stat1 <- TukeyHSD(res.aov.enha.HFDup.stat)
res.aov.enha.HFDup.stat2 <- res.aov.enha.HFDup.stat1$name

counts_enha_HFDdown <- counts_enha_norm3 %>% filter(loc_ID%in%Enha_Diffbind_HFDdown$loc_ID) %>% dplyr::select(1,20:25) 
counts_enha_HFDdown2 <- counts_enha_HFDdown %>%
  pivot_longer(cols=2:7) %>% mutate("updown"="HFDdown")

# STATISTICS - One-sided anova.
res.aov.enha.HFDdown.stat <- aov(value ~ name, data = counts_enha_HFDdown2)
summary(res.aov.enha.HFDdown.stat)
res.aov.enha.HFDdown.stat1 <- TukeyHSD(res.aov.enha.HFDdown.stat)
res.aov.enha.HFDdown.stat2 <- res.aov.enha.HFDdown.stat1$name

counts_enha_HFD_plot <- rbind(counts_enha_HFDup2, counts_enha_HFDdown2)
counts_enha_HFD_plot$value <- counts_enha_HFD_plot$value+1
```
#### Plot for enhancers
```{r}
ggplot(counts_enha_HFD_plot, aes(x=factor(name, levels=order), y=log2(value), fill=updown)) +
  geom_boxplot(alpha=0.8) +
  theme_classic() +
  facet_wrap(vars(updown)) + 
  xlab("") +
  ggtitle("Enhancers") +
  ylab("H3K27ac signal\n log2(normalized read count in \nDAc enhancer peaks in CDvsHFD)") +
   scale_fill_manual(values = c("#88CCEE",  "#332288", "#DDCC77", "#CC6677", "#AA4499",  "#882255"))
```

```{r}
sessionInfo()
```

